---
title: "Time Series Analysis and Forecasting. Tutorial 1"
output: html_notebook
---

This notebook is based on the examples from the Time Series Analysis and Forecasting lectures by Yarema Okhrin.

The objective is to forecast the credit score of bank clients relying on several explanatory variables.

```{r}
library(fpp)
library(GGally)
library(forecast)
library(Metrics)
library(margins)
```

### 1. Import the data and make appropriate visualisation (pairs, ggpairs)
```{r}
data(credit)
```

```{r}
pairs(credit)
```

```{r results='hide',fig.keep='all'}
ggpairs(credit)
```

### 2. Split the data randomly into training and test data set (sample)
```{r}
# Sample indexes
indexes <- sample(1:nrow(credit), size=0.2*nrow(credit))

# Split the data
credit.test <- credit[indexes,]
dim(credit.test)  # 100 8

credit.train <- credit[-indexes,]
dim(credit.train) # 400 11
```

### 3. Run the regression using training data set. Assess the fit
```{r}
model.simple <- lm(score ~ ., credit.train)
summary(model.simple)
plot(model.simple)
```

### 4. Compute the confidence intervals for the parameters.
```{r}
confint(model.simple)
```

### 5. Compute the forecasts (point, prediction and confidence intervals) for the test data set (predict, forecast)
```{r}
pred <- predict(model.simple, newdata = credit.test, interval = "confidence")
matplot(c(1:100), pred, type="p", pch=15, col=c(1,2,2), cex=0.8)
pred <- predict(model.simple, newdata = credit.test, interval = "prediction")
forec <- forecast(model.simple, newdata = credit.test)
```

### 6. Compute forecast errors and analyse them
```{r}
mse(credit.test$score, forec$mean)
mae(credit.test$score, forec$mean)
mape(credit.test$score, forec$mean)
```

```{r}
# Plot forecast errors
plot(credit.test$score, credit.test$score - forec$mean)
```

### 7. Make appropriate data transformation and repeat the above steps for the new model
```{r}
model.transformed <-
  lm(score ~ savings + I(income ^ 2) + log(savings + 1),
  data = credit.train)

summary(model.transformed)
```

```{r}
confint(model.transformed)
```

```{r}
# Model with logs
col.names <- names(credit)#[2:ncol(credit)]

credit.test.log = cbind(
  credit.test[,1], log(credit.test[,2:3] + 1),
  credit.test[,4:5], log(credit.test[,6:7] + 1))

names(credit.test.log) <- col.names

credit.train.log = cbind(
  credit.train[,1], log(credit.train[,2:3]+1),
  credit.train[,4:5], log(credit.train[,6:7]+1))

names(credit.train.log) <- col.names
```

```{r}
# Model and forecasts
model.log = lm(score ~ ., credit.train.log)

pred.log = predict(
  model.log,
  newdata = credit.test.log,
  interval = "confidence")

forec.log = forecast(model.log, newdata = credit.test.log)
```

```{r}
# Plot forecast errors
plot(credit.test.log$score, credit.test.log$score - forec.log$mean)
```

### 8. Compute the marginal effects for the original and the transformed models (margins)
```{r}
model <- lm(score ~ savings + income, credit.train)
margins(model)
```

### 9. Compare the forecasting ability of the two models using loss functions and using appropriate tests

### 10. The score variable is slightly skewed. Takings logs shall overcome this problem. Check the forecasting ability of the log-score model
