---
title: "Time Series Analysis and Forecasting. Final Assignment"
output:
  html_notebook: default
  pdf_document: default
---

```{r}
library(xts)
library(forecast)
library(car)
library(ggplot2)
```

### 1. Regression-type modelling and forecasting

**(a) Import the 1st data set to R. Create a ts-object. Keep the last 5 years of data as test data set.**

```{r}
beerwine <- read.csv('../data/BeerWineUS.csv')
beerwine <- ts(beerwine$MRTSSM4453USN, start = 1992, frequency = 12)

N <- length(beerwine)
beerwine.train <- window(beerwine, end=2012.99)
beerwine.test <- window(beerwine, start=2013)

print(beerwine.test)
```

**(b) Plot the data and choose an appropriate order of time polynomial for modelling. Run a simple regression on the time polynomial and check the autocorrelation of the residuals (also using Durbin-Watson). What can be concluded?**

```{r}
plot(beerwine.train)
```

```{r}
t <- time(beerwine.train)
fit <- lm(beerwine.train ~ I(t^2) + I(t))
prd <- data.frame(t = t)
err <- predict(fit, newdata = prd, se.fit = TRUE)

prd$lci <- err$fit - 1.96 * err$se.fit
prd$fit <- err$fit
prd$uci <- err$fit + 1.96 * err$se.fit

ggplot(prd, aes(x = t, y = fit)) +
  theme_bw() +
  geom_line() +
  geom_smooth(aes(ymin = lci, ymax = uci), stat = "identity") +
  geom_line(data = beerwine.train, aes(x = t, y = beerwine.train))
```

```{r}
summary(fit)
```

```{r}
durbinWatsonTest(fit)
```

p-value is pretty high, which means that we should reject the null hypothesis about autocorrelation beign zero. This means that the residuals are autocorrelated.

```{r}
acf(fit$residuals)
```

**(c) Since the time series is rather short to define a dummy for every month, it is rea- sonable to define dummies for longer periods. Use ggmonthplot or ggseasonplot to decide if it is possible to assign dummies to quarters/specific months, etc. Run a corresponding regression and check the residuals as above.**

```{r}
ggmonthplot(beerwine.train)
```

```{r}
ggseasonplot(beerwine.train)
```

```{r}
m <- cycle(beerwine.train)
Q1 <- as.numeric(m >= 1 & m <= 7)
Q2 <- as.numeric(m >= 8 & m <= 11)
Q3 <- as.numeric(m == 12)
```

```{r}
t <- time(beerwine.train)
t1 <- I(t)
t2 <- I(t^2)
beerwine.train.df <- as.data.frame(cbind(beerwine.train, t1, t2, Q1, Q2, Q3))
```

```{r}
fit <- lm(beerwine.train ~ ., data = beerwine.train.df)
```

```{r}
summary(fit)
```

```{r warning=FALSE}
t <- time(beerwine.train)
prd <- data.frame(t = t)
err <- predict(fit, newdata = prd, se.fit = TRUE)

prd$lci <- err$fit - 1.96 * err$se.fit
prd$fit <- err$fit
prd$uci <- err$fit + 1.96 * err$se.fit

ggplot(prd, aes(x = t, y = fit)) +
  theme_bw() +
  geom_line() +
  geom_smooth(aes(ymin = lci, ymax = uci), stat = "identity") +
  geom_line(data = beerwine.train, aes(x = t, y = beerwine.train))
```

```{r}
durbinWatsonTest(fit)
```

```{r}
acf(fit$residuals)
```

**(d) With the last model compute the (interval) forecasts for the test data set and visualize the results.**

```{r}
m <- cycle(beerwine.test)
Q1 <- as.numeric(m >= 1 & m <= 7)
Q2 <- as.numeric(m >= 8 & m <= 11)
Q3 <- as.numeric(m == 12)
```

```{r}
t <- time(beerwine.test)
t1 <- I(t)
t2 <- I(t^2)
beerwine.test.df <- as.data.frame(cbind(beerwine.test, t1, t2, Q1, Q2, Q3))
```

```{r}
pred <- as.ts(cbind(beerwine.test, predict(fit, newdata=beerwine.test.df)))
res = as.ts(cbind(beerwine.train, fit$fitted))
```

```{r}
plot(pred[,2])
```

### 2. Time series decomposition

**(a) Try several moving average techniques to extract the trend (ma). Which orders/form of moving averages do provide the best results?**

```{r}
tau <- 10;

get.maf <- function(width) {
  function(series) {
    rollapply(series, width = width, by = 1, FUN = mean, align = "right")
  }
}
```

```{r}
maf3 <- get.maf(3)
maf5 <- get.maf(5)
maf7 <- get.maf(7)
```

```{r}
s <- beerwine.train
all.forec <- cbind(maf3(s), maf5(s), maf7(s))
all <- cbind(s, all.forec);
plot.ts(all, plot.type="single", col=c(1:6,1:5))

model.names <- c("true", "maf 3", "maf 5", "maf 7")
legend(x = "bottomright", legend=model.names, ncol=3, bty="n", col=c(1:6,1:5), lty=c(rep(1,6),rep(2,6)))
```

**(b) Apply the loess regression for trend extraction.**

We apply R's stl() function ("seasonal and trend decomposition using Loess") to decompose beerwine data

```{r}
decomposed <- stl(beerwine.train, s.window="periodic")
trend <- decomposed$time.series[,2]
plot(trend)
```

**(c) For one of the above computed trends compute the seasonal and the irregular components. Visualize the decomposition.**

```{r}
seasonal <- decomposed$time.series[,1]
irregular <- decomposed$time.series[,3]
```

```{r}
plot(decomposed)
```

**(d) Check the ACF of the irregular component. Is there anything left in the time series?**

```{r}
acf(irregular)
```

Yes, there are big spikes at lags 1 and 2. So the current value still depends consistently on the value of the same month in a previous year.

### 3. Simple forecasting and forecasting with exponential smoothing

**(a) Compute simple one-step-ahead forecasts for the test data set using naıve, absolute trend and relative trend methods. Compute the corresponding losses.**

```{r}
tau <- length(beerwine.test) - 1

seas.naive <- function(series) {
  sapply(tau:1, function(x) snaive(head(series, -x), 1)$mean)
}

abs.trend <- function(series) {
  n <- length(series)
  2 * series[(n - tau):(n - 1)] - series[(n - tau - 1):(n - 2)]
}

rel.trend <- function(series) {
  n <- length(series)
  series[(n - tau):(n - 1)] ^ 2 / series[(n - tau - 1):(n - 2)]
}
```

```{r warning=FALSE}
s <- beerwine.test
all.forec <- cbind(seas.naive(s), abs.trend(s), rel.trend(s))
all <- cbind(tail(s, tau), all.forec)
plot.ts(all, plot.type="single", col=c(1:6,1:5))

model.names <- c("true", "naive", "absolute trend", "relative trend")
legend(x = "topleft", legend=model.names, ncol=3, bty="n", col=c(1:6,1:5), lty=c(rep(1,6),rep(2,6)))
```

```{r}
loss.functions <- function(x.hat, x) {
  c(mean((x-x.hat)^2), mean(abs(x-x.hat)), mean(abs( (x-x.hat)/x )) )
}
```

```{r}
apply(all.forec, 2, function(x) loss.functions(x, tail(beerwine.test, tau)))
```