---
title: "Time Series Analysis and Forecasting. Tutorial 4"
output: html_notebook
---

This notebook is based on the examples from the Time Series Analysis and Forecasting lectures by Yarema Okhrin.

```{r}
library("forecast")
library("fpp")
```

### Loading the dataset
We will be using [elecequip](https://www.rdocumentation.org/packages/fpp/versions/0.5/topics/elecequip) - a time series of Electrical Equipment Manufactured In The Euro Area. This dataset is loaded from [fpp](https://www.rdocumentation.org/packages/fpp/versions/0.5) package. Data is adjusted by working days; Euro area (16 countries).

```{r}
data(elecequip)
head(elecequip, n = 48)
```

```{r}
# Constants
tau <- 10;
```

### Defining forecasting functions
```{r}
naive.forec <- function(series) {
  sapply(tau:1, function(x) naive(head(series, -x), 1)$mean)
}

seas.naive <- function(series) {
  sapply(tau:1, function(x) snaive(head(series, -x), 1)$mean)
}

abs.trend <- function(series) {
  n <- length(series)
  2 * series[(n - tau):(n - 1)] - series[(n - tau - 1):(n - 2)]
}

rel.trend <- function(series) {
  n <- length(series)
  series[(n - tau):(n - 1)] ^ 2 / series[(n - tau - 1):(n - 2)]
}
```

```{r}
get.maf <- function(width) {
  function(series) {
    tail(rollapply(series, width = width, by = 1, FUN = mean, align = "right"), tau)
  }
}

get.holtwinters <- function(beta = NULL, gamma = NULL) {
  function(series) {
    my.mod = HoltWinters(series, beta=beta, gamma=gamma)
    tail(my.mod$fitted[,1], tau)
  }
}
```

```{r}
forec.funcs <- c(
  naive.forec,
  seas.naive,
  abs.trend,
  rel.trend,
  get.maf(width = 3),
  get.maf(width = 5),
  get.maf(width = 7),
  get.holtwinters(beta = FALSE, gamma = FALSE),
  get.holtwinters(gamma = FALSE),
  get.holtwinters()
);
```

### Forecasting
```{r}
all.forec <- matrix(0, ncol=10, nrow=tau);
for(i in 1:10) {
  all.forec[,i] <- forec.funcs[[i]](elecequip);
}
```

### Computing losses
```{r}
loss.functions <- function(x.hat, x) {
  c(mean((x-x.hat)^2), mean(abs(x-x.hat)), mean(abs( (x-x.hat)/x )) )
}
```

```{r}
apply(all.forec, 2, function(x) loss.functions(x, tail(elecequip,tau)))
```

### Plotting forecasts
```{r}
# plot all forecasts
model.names <- c("true", "naive", "seas. naive","abs trend", "rel. trend", "maf 3", "maf 5", "maf 7", "emwa", "holt", "holtwinters")
all <- cbind(tail(elecequip,tau), all.forec);
names(all) <- model.names;
plot.ts(all, plot.type="single", col=c(1:6,1:5), lty=c(rep(1,6),rep(2,6)), ylab="forecasts", ylim=c(30,120))
legend(x = "bottomleft", legend=model.names, ncol=3, bty="n", col=c(1:6,1:5), lty=c(rep(1,6),rep(2,6)))
```

```{r}
# plot holtwinters
hw <- HoltWinters(elecequip);
plot(hw$fitted)
```